Estimating the association between Facebook adoption and well-being

### 2022-09-16 ms.Rmd
edits

abstract: |
  Social media's potential effects on well-being have received considerable research interest, but much of past work is hampered by an exclusive focus on demographics in the Global North and inaccurate self-reports of social media engagement. We describe associations linking 72 countries' Facebook adoption to the well-being of 946,798 individuals from 2008 to 2019. We found no associations suggesting that the global penetration of social media has caused widespread psychological harm: Facebook adoption predicted life satisfaction and positive experiences positively, and negative experiences negatively, both between countries and within countries but over time. Facebook adoption was more positively associated with well-being for younger individuals, but country-specific results were mixed. In order to move beyond studying aggregates and to better understand social media's causal effects, we need more transparent collaborative research between independent scientists and the technology industry.

significance: |
  Billions of individuals across the globe use social media on a regular basis. The widespread adoption of social media---most prominently Facebook---has fundamentally shifted how humans conduct their lives. Consequently, researchers, healthcare professionals, and policy makers have repeatedly raised alarms that the spread of social media may be negatively affecting individuals, particularly young ones, mental health and well-being on a global scale. We examined the plausibility of this idea by associating 72 nations' yearly Facebook adoption rates to the well-being of 946,798 individuals from 2008 to 2019. Overall, we found that increased rates of daily and monthly active Facebook users predicted greater levels of well-being, when comparing countries to other countries, or countries to themselves but over time. These findings should guide debate surrounding social media towards more empirical foundations: The importance of this topic requires that independent scientists gain access to more detailed data from the technology industry in order to better determine how and why modern online platforms might be affecting their users.

@@

@@

The ways in which people use technology for most domains in life has changed dramatically since the mass introduction of the internet in the 1990s, and the subsequent technologies facilitated by it. Most prominently, the popularization of modern social media platforms in 2008 precipitated widespread changes to human activities via features such as marketplaces, personalized news feeds, photo sharing, live streaming, and other features that the "metaverse" promises to build on. The first social media with global adoption, MySpace (launched 2003) saw 115 million users in 2008---the year in which it was replaced as the leading platform by Facebook (2004). Today (2022-Q1), Facebook reports 2.94 billion monthly active users [@metaMetaReportsFirst2022], or about one third of the global population. Along with social media's global penetration, debate surrounding their potential effects on individual and collective well-being has intensified.

Although reports of negative psychological outcomes associated social media are common in academic writing [@krossFacebookUsePredicts2013] and popular media [@thompsonSocialMediaAttention2021], evidence for harms is, on balance, more speculative than conclusive [@bestOnlineCommunicationSocial2014; @dickson2018screen; @odgersAnnualResearchReview2020]. Recent results on the associations between social media use and well-being are mixed and depend on arbitrary analytic choices [@orbenSocialMediaEnduring2019]. Other studies have reported that there have been few if any changes in associations linking technology use to mental health in this period of social media's global adoption [@vuorreThereNoEvidence2021]. A general lack of validated measures, poorly specified causal models, and inadequate data have yielded a large number of low-quality studies [@parrySystematicReviewMetaanalysis2021; @sewallDoesObjectivelyMeasured2022]. Furthermore, because nearly all investigations have focused on samples from the Global North [@ghaiSocialMediaAdolescent2022], we have next to no idea of how the wide adoption of social media platforms relates to psychological well-being across the world.

To better understand the plausible range of associations, we linked data tracking Facebook's global adoption with three indicators of global well-being. We examined 72 countries' per capita daily active Facebook users (DAU) in two age brackets (13-34 and 35+ years) as predictors of life satisfaction (LS), negative (NE), and positive psychological experiences (PE) at the level of years spanning 2008 to 2019. The well-being data represented 946,798 individuals' responses from the nationally representative Gallup World Poll Survey [@gallupHowDoesGallup2014]. We joined these unique datasets to conduct a descriptive study to answer three basic yet important questions: First, to what extent did Facebook adoption predict well-being? Second, do these associations differ by age or sex. And finally, how might these associations have differed between countries?

@@

@@

# Figure of all variables over time
p_a <- d_fb %>% 
  ggplot(aes(year, dau, group = country)) +
  scale_y_continuous(
    "Percentage of population", 
    expand = expansion(.03),
    breaks = extended_breaks(),
    labels = ~percent(.)
  ) +
  scale_x_continuous(
    "Year", 
    breaks = seq(2008, 2018, by = 2), 
    labels = ~str_replace(., "20", "'"),
    expand = expansion(0.01)
  ) +
  geom_line(size = .125) +
  # geom_smooth(group = 1, method = "gam", se = F) +
  facet_wrap("age", nrow = 1) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  theme(text = element_text(size = 5))

p_b <- d %>% 
  # Exclude rows where FB data existed but not GWP data
  drop_na(val) %>% 
  # Actual lowest age in GWP
  mutate(age = factor(age, labels = c("15-34", "35+"))) %>% 
  mutate(
    outcome = case_when(
      outcome == "Life_satisfaction" ~ "LS",
      outcome == "Negative_experiences" ~ "NE",
      outcome == "Positive_experiences" ~ "PE"
    )
  ) %>% 
  mutate(panel = str_glue("{age} ({sex})")) %>% 
  ggplot(aes(year, val, group = country)) +
  scale_color_brewer(palette = "Dark2") +
  scale_y_continuous(
    "Value", 
    expand = expansion(.01),
    breaks = extended_breaks(),
    labels = ~percent(.)
  ) +
  scale_x_continuous(
    "Year", 
    breaks = seq(2008, 2018, by = 2),
    labels = ~str_replace(., "20", "'")
  ) +
  geom_line(size = .1) +
  facet_grid(
    outcome~panel, 
    scales = "free_y",
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  theme(legend.position = "none") +
  theme(text = element_text(size = 5))

@@

To answer our first question of how Facebook adoption relates to well-being, we specified a multilevel linear meta-regression model of each well-being outcome on time, sex, age, and DAU (both within- and between-country-centered, see SI). We included all interactions between sex, age, and time, and sex, age, and DAU (within-country-centered), and allowed all parameters to vary randomly across countries. We first examined whether the relative standing of countries on their average Facebook adoption predicted well-being (the model's between-country associations). We found that countries with greater average DAU had somewhat higher levels of life satisfaction and positive experiences, and lower levels of negative experiences than countries with lower DAU (Table 1). However, there are large and important differences between countries in factors that might underlie differences in both social media adoption and well-being, such as socioeconomic conditions. As such, while descriptively informative, these associations are likely to indicate between-country confounding factors.

@@

tmp %>%
  ggplot() +
  aes(
    x = Estimate, 
    y = fct_rev(country), 
    col = Age, 
    fill = interaction(Age, Star, sep = " "), 
    shape = Sex
  ) +
  scale_y_discrete(
    "Country",
    expand = expansion(.015),
    labels = ~str_replace(., "World", "Average")
  ) +
  scale_x_continuous(
    "Association with daily active Facebook users (95%CI)",
    breaks = extended_breaks(7),
    expand = expansion(.1)
  ) +
  scale_shape_manual(
    values = c(21, 22)
  ) +
  scale_color_brewer(
    "Age", palette = "Set1", aesthetics = c("color")
  ) +
  scale_fill_manual(values = c("white", "white", "#E41A1C", "#377EB8")) +
  geom_vline(xintercept = 0, lty = 2, size = .25, col = "grey50") +
  geom_linerangeh(
    aes(xmin = CI.Lower, xmax = CI.Upper),
    size = .15, position = position_dodge2v(.7), 
    show.legend = FALSE
  ) +
  geom_point(
    size = .65, stroke = .45,
    position = position_dodge2v(.7)
  ) +
  facet_wrap(
    "outcome", scales = "free_x", nrow = 1,
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(size = 1.5)),
    shape = guide_legend(override.aes = list(size = 1.5))
  ) +
  theme(
    text = element_text(size = 5),
    axis.text.y = element_text(size = rel(0.9)),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(
      size = .1, linetype = 2, color = "gray60"
    ),
    legend.position = "bottom", 
    legend.box.margin = margin(-15, 0, -10, 0, "pt"),
    legend.background = element_rect(fill = NA),
    legend.key = element_rect(fill = NA),
    legend.text = element_text(size = rel(0.9))
  )
```

Variations in such confounding factors are likely to be significantly smaller within countries over time. We therefore next focused on the model's within-country associations, which measure the extent to which Facebook adoption in a given country predicted well-being in that country adjusting for general linear temporal trends in the well-being outcome. On average across countries, age, and sex, we found that years with greater than the average country's average DAU were associated with greater levels of life satisfaction and positive experiences, and lower levels of negative experiences (Table 1). For the average country in our sample, Facebook adoption positively predicted well-being. While these associations indicate relations within countries and adjust for confounders that vary linearly with time, by including time as a predictor, they are still susceptible to confounders and do not indicate causal relations. Rather, this association describes the fact that, all else being equal, years with greater Facebook adoption tended to be those with greater levels of well-being.

We then turned to our second question and assessed whether within-country associations linking Facebook adoption to well-being differed between the demographic groups. Differences in the association linking Facebook adoption and well-being were greater between the younger and older demographics than they were between females and males (Figure 2, Average). DAU predicted life satisfaction more positively in the younger age bracket, but these estimates' CIs did not exclude zero. DAU predicted negative experiences negatively in the younger age group, but positively for older males. These differences were less pronounced for positive experiences, which were positively associated with DAU only in the 13-34 year old male group. These results show that Facebook adoption was, for the average nation in our sample, more positively predictive of well-being for individuals below 35, and that differences between males and females were generally smaller than those between age groups.

However, while informative aggregates, these results do not describe associations between Facebook adoption and well-being for any individual country, but rather their averages. To answer our third question, we computed country-specific estimates for each sex and age group (Figure 2). On average, for life satisfaction, `r tmp[1, 5]` countries credibly positive associations. For negative experiences, `r tmp[2, 6]` countries had a credibly negative association, and `r tmp[3, 5]` countries had a credibly positive association between DAU and positive experiences. There were no credible (at the 95% level) country-level negative associations between DAU and life-satisfaction. Although many country-level estimates were uncertain, these results aligned with the general pattern of positive associations between Facebook adoption and country-level well-being.

```{r fig-ce-prepare}
#| eval: false
# Create a predictor table with evenly spaced mau_cw values for each country
newx_country <- fits$data[[1]] %>% 
  group_by(country, age, sex, mau_cb) %>% 
  filter(mau_cw %in% c(min(mau_cw), max(mau_cw))) %>% 
  select(country, age, sex, mau_cb, mau_cw) %>% 
  # For above groupings, create a sequence of dau_cws from min to max
  group_modify(
    ~tibble(mau_cw = seq(min(.$mau_cw), max(.$mau_cw), length = 31))
  ) %>% 
  ungroup() %>% 
  mutate(year = 0, se = 0)

newx_world <- fits$data[[1]] %>% 
  group_by(age, sex) %>% 
  filter(mau_cw %in% c(min(mau_cw), max(mau_cw))) %>% 
  select(age, sex, mau_cw) %>% 
  # For above groupings, create a sequence of mau_cws from min to max
  group_modify(
    ~tibble(mau_cw = seq(min(.$mau_cw), max(.$mau_cw), length = 31))
  ) %>% 
  ungroup() %>% 
  mutate(mau_cb = 0, year = 0, se = 0)

newy <- fits %>% 
  mutate(
    country = map(
      mau, 
      ~bind_cols(newx_country, fitted(.x, newx_country))
    ),
    world = map(
      mau, 
      ~bind_cols(newx_world, fitted(.x, newx_world, re_formula = NA))
    )
  ) %>% 
  select(outcome, country, world)
```

```{r fig-ce}
#| eval: false
#| include: true

newy %>% 
  pivot_longer(c(country, world)) %>% 
  unnest(value) %>% 
  ungroup() %>% 
  mutate(country = if_else(name == "world", "World", country)) %>% 
  ggplot(
    aes(mau_cw, Estimate, group = interaction(country, sex), col = sex, fill = sex)
  ) +
  scale_color_brewer(
    "Age", 
    palette = "Dark2", 
    aesthetics = c("color", "fill")
  ) +
  scale_y_continuous(
    "Estimated outcome",
    breaks = extended_breaks()
  ) +
  scale_x_continuous(
    "Country-mean deviated monthly FB users",
    breaks = extended_breaks(),
    labels = ~percent(.)
  ) +
  scale_size_manual(values = c(.2, .4)) +
  geom_ribbon(
    data = . %>% filter(name == "world"),
    aes(ymin = Q2.5, ymax = Q97.5),
    alpha = .2, col = NA
  ) +
  geom_line(aes(size = name)) +
  facet_grid(
    age~outcome, scales = "free_y",
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  theme(legend.position = "none")
```

We found that, overall, Facebook adoption positively predicted well-being. This association held when comparing countries to other countries and to themselves over time. While these descriptive results do not speak to causal effects, they align with other recent findings suggesting that technology has not become increasingly harmful over time [@vuorreThereNoEvidence2021], and that increased adoption of general internet technologies are not, overall, associated with psychological harms [@vuorreGlobalWellbeingMental2022]. That said, these conclusions are qualified because our focus on daily active users could miss those who use Facebook less regularly. To test this possibility, we also conducted our analyses using monthly active Facebook users (MAU) as the predictor. Those results were substantively the same (see online supplement).

@@

Both datasets (Facebook and Gallup) are proprietary and we therefore could not share them with this manuscript.  The analytic code, along with synthetic datasets, is available at (LINK). Researchers can contact Facebook (ccobb@fb.com) to access our code, the Facebook data, and a synthetic version of the Gallup data for reproducing and extending our analyses. The Gallup data are available online to subscribing institutions.

@@

We studied two metrics of Facebook adoption, daily (DAU) and monthly active users (MAU), from 2008 to 2019 for 72 countries (see Figure 2). FB estimates yearly DAU & MAU with their corresponding means from June 1 to August 31, and rounds values greater than 10,000,000 to three significant digits, and values less than 10,000,000 to the nearest 10,000. We report analyses based on DAU in the main manuscript, and MAU in the online supplement. DAU and MAU estimates are calculated separately for individuals aged 13-34 and 35+. User age is determined based on Facebook profile information, which can be inaccurate (e.g. young individuals report an older age.) Accordingly, 0.008% of total MAU were trimmed to exclude accounts with unrealistic on non-reported age.

@@

Facebook defines users as individuals who visited Facebook or used Messenger during the appropriate period. Although accurately captured, these numbers are not perfect indicators of actual user numbers because of possible duplicate and false accounts. FB estimates those accounts to count for 11% and 5% of global MAUs, and that the former may be more likely in developing regions. Because internal criteria and methodology for determining duplicate and false accounts can change over time, estimates of MAU and DAU can also change.

@@

$$
\begin{aligned} 
y_i &\sim \text{Normal}(\mu_i, \sigma^{y2}v_i), \\
\mu_i &= \alpha_{0} + \beta_{0\text{country}[i]} + \\
&\quad\ (\alpha_{1} + \beta_{1\text{country}[i]})\text{Sex}_i + \\
&\quad\ (\alpha_{2} + \beta_{2\text{country}[i]})\text{Age}_i + \\
&\quad\ (\alpha_{3} + \beta_{3\text{country}[i]})\text{Time}_i + \\
&\quad\ (\alpha_{4} + \beta_{4\text{country}[i]})\text{DAU}^{\text{CW}}_i + \\
&\quad\ (\alpha_{5} + \beta_{5\text{country}[i]})\text{Sex}_i \times \text{Time}_i + \\
&\quad\ (\alpha_{6} + \beta_{6\text{country}[i]})\text{Age}_i \times \text{Time}_i + \\
&\quad\ (\alpha_{7} + \beta_{7\text{country}[i]})\text{Sex}_i \times \text{Age}_i + \\
&\quad\ (\alpha_{8} + \beta_{8\text{country}[i]})\text{Sex}_i \times \text{Age}_i \times \text{Time}_i + \\
&\quad\ (\alpha_{9} + \beta_{9\text{country}[i]})\text{Sex}_i \times \text{DAU}^{\text{CW}}_i + \\
&\quad\ (\alpha_{10} + \beta_{10\text{country}[i]})\text{Age}_i \times \text{DAU}^{\text{CW}}_i + \\
&\quad\ (\alpha_{11} + \beta_{11\text{country}[i]})\text{Sex}_i \times \text{Age}_i \times \text{DAU}^{\text{CW}}_i + \\
&\quad\ \alpha_{12} \text{DAU}^{\text{CB}}_i,
\\
\pmb{\beta} &\sim \text{MVN}(\pmb{0}, \Sigma).
\end{aligned}
$$

Where $y_i$ is the outcome (e.g. life satisfaction) on row $i$, $v_i$ its sampling variance, $\alpha$ are the population-level coefficients, $\beta_{0, \dots, 11\text{country}[i]}$ are the country-specific coefficients for the country indicated on row $i$, $\text{DAU}^{\text{CW}}_i$ the within-country centered year-aggregated daily (or monthly) active Facebook users, and $\text{DAU}^{\text{CB}}_i$ the between-country centered DAU (or MAU).

@@

tmp %>% 
  ggplot() +
  aes(
    x = Estimate, 
    y = fct_rev(country), 
    col = Age, 
    fill = interaction(Age, Star, sep = " "), 
    shape = Sex
  ) +
  scale_y_discrete(
    "Country",
    expand = expansion(.015),
    labels = ~str_replace(., "World", "Average")
  ) +
  scale_x_continuous(
    "Association with monthly active Facebook users (95%CI)",
    breaks = extended_breaks(7),
    expand = expansion(.1)
  ) +
  scale_shape_manual(
    values = c(21, 22)
  ) +
  scale_color_brewer(
    "Age", palette = "Set1", aesthetics = c("color")
  ) +
  scale_fill_manual(values = c("white", "white", "#E41A1C", "#377EB8")) +
  geom_vline(xintercept = 0, lty = 2, size = .25, col = "grey50") +
  geom_linerangeh(
    aes(xmin = CI.Lower, xmax = CI.Upper),
    size = .15, position = position_dodge2v(.7), 
    show.legend = FALSE
  ) +
  geom_point(
    size = .65, stroke = .45,
    position = position_dodge2v(.7)
  ) +
  facet_wrap(
    "outcome", scales = "free_x", nrow = 1,
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(size = 1.5)),
    shape = guide_legend(override.aes = list(size = 1.5))
  ) +
  theme(
    text = element_text(size = 5),
    axis.text.y = element_text(size = rel(0.9)),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(
      size = .1, linetype = 2, color = "gray60"
    ),
    legend.position = "bottom", 
    legend.box.margin = margin(-15, 0, -10, 0, "pt"),
    legend.background = element_rect(fill = NA),
    legend.key = element_rect(fill = NA),
    legend.text = element_text(size = rel(0.9))
  )
```

>>>>

abstract: |
  Social media's potential effects on well-being have received considerable research interest, but much of past work is hampered by an exclusive focus on demographics in the Global North and inaccurate self-reports of social media engagement. We describe associations linking 72 countries' Facebook adoption to the well-being of 946,798 individuals from 2008 to 2019. We found no associations suggesting that the global penetration of social media has caused widespread psychological harm: Facebook adoption predicted life satisfaction and positive experiences positively, and negative experiences negatively, both between countries and within countries but over time, and the associations were small. Facebook adoption was more positively associated with well-being for younger individuals, but country-specific results were mixed. In order to move beyond studying aggregates and to better understand social media's causal effects, we need more transparent collaborative research between independent scientists and the technology industry.

significance: |
  Billions of individuals across the globe use social media on a regular basis. The widespread adoption of social media---most prominently Facebook---has fundamentally shifted how humans conduct their lives. Consequently, researchers, healthcare professionals, and policy makers have repeatedly raised alarms that the spread of social media may be negatively affecting individuals', particularly young ones', mental health and well-being on a global scale. We examined the plausibility of this idea by associating 72 nations' yearly Facebook adoption rates to the well-being of 946,798 individuals from 2008 to 2019. Overall, we found that increased rates of daily and monthly active Facebook users slightly predicted greater levels of well-being, when comparing countries to other countries, or countries to themselves but over time. These findings should guide debate surrounding social media towards more empirical foundations: The importance of this topic requires that independent scientists gain access to more detailed data from the technology industry in order to better determine how, when, and why modern online platforms might be affecting their users.

@@

\raggedbottom

@@

The ways in which people use technology for most domains in life has changed dramatically since the mass introduction of the Internet in the 1990s, and the subsequent technologies facilitated by it. Most prominently, the popularization of modern social media platforms circa 2008 precipitated widespread changes to human activities via features such as marketplaces, personalized news feeds, photo sharing, live streaming, and other features that the "metaverse" now promises to build on. The first social media with broad adoption, MySpace (launched 2003) saw 115 million users in 2008---the year in which it was replaced as the leading platform by Facebook (2004). Today (2022-Q1), Facebook reports 2.94 billion monthly active users [@metaMetaReportsFirst2022], or about one third of the global population. Along with social media's global penetration, debate surrounding their potential effects on individual and collective well-being has intensified.

Although reports of negative psychological outcomes associated with social media are common in academic and popular writing [@krossFacebookUsePredicts2013; @thompsonSocialMediaAttention2021], evidence for harms is, on balance, more speculative than conclusive [@bestOnlineCommunicationSocial2014; @dickson2018screen; @odgersAnnualResearchReview2020]. Recent results on the associations between social media use and well-being are mixed and depend on arbitrary analytic choices [@orbenSocialMediaEnduring2019]. Other studies have reported that there have been few if any changes in associations linking technology use to mental health in this period of social media's global adoption [@vuorreThereNoEvidence2021]. A general lack of validated measures, poorly specified causal models, and inadequate data have yielded a large number of low-quality studies [@parrySystematicReviewMetaanalysis2021; @sewallDoesObjectivelyMeasured2022]. Furthermore, because nearly all investigations have focused on samples from the Global North [@ghaiSocialMediaAdolescent2022], we have next to no idea of how the wider adoption of social media platforms relates to psychological well-being across the world.

To better understand the plausible range of associations, we linked data tracking Facebook's global adoption with three indicators of well-being. We examined 72 countries' per capita daily active Facebook users (DAU) in two age brackets (13-34 and 35+ years) as predictors of life satisfaction (LS), negative (NE), and positive psychological experiences (PE) at the level of years spanning 2008 to 2019. The well-being data represented 946,798 individuals' responses from the nationally representative Gallup World Poll Survey [@gallupHowDoesGallup2014]. We joined these unique datasets to conduct a descriptive study to answer three basic yet important questions: First, to what extent is Facebook adoption associated with well-being? Second, do these associations differ by age or sex. And finally, how might these associations have differed between countries?

@@

To answer our first question of how Facebook adoption relates to well-being, we specified a multilevel linear meta-regression model of each well-being outcome on time, sex, age, and DAU (both within- and between-country-centered, see SI). We included all interactions between sex, age, and time, and sex, age, and DAU (within-country-centered), and allowed all parameters to vary randomly across countries. We first examined whether the relative standing of countries on their average Facebook adoption predicted well-being (the model's between-country associations). We found that countries with greater average DAU had somewhat higher levels of life satisfaction and positive experiences, and lower levels of negative experiences than countries with lower DAU (Table 1). However, there are large and important differences between countries in factors that might underlie differences in both social media adoption and well-being, such as socioeconomic conditions. While descriptively informative, these associations are likely to indicate between-country confounding factors.

@@

# Figure of all variables over time
p_a <- d_fb %>% 
  ggplot(aes(year, dau, group = country)) +
  scale_y_continuous(
    "Percentage of population", 
    expand = expansion(.03),
    breaks = extended_breaks(),
    labels = ~percent(.)
  ) +
  scale_x_continuous(
    "Year", 
    breaks = seq(2008, 2018, by = 2), 
    labels = ~str_replace(., "20", "'"),
    expand = expansion(0.01)
  ) +
  geom_line(size = .125) +
  # geom_smooth(group = 1, method = "gam", se = F) +
  facet_wrap("age", nrow = 1) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  theme(
    strip.text.x = element_text(margin = margin(2,0,2,0, "pt")),
    text = element_text(size = 5)
    )

p_b <- d %>% 
  # Exclude rows where FB data existed but not GWP data
  drop_na(val) %>% 
  # Actual lowest age in GWP
  mutate(age = factor(age, labels = c("15-34", "35+"))) %>% 
  mutate(
    outcome = case_when(
      outcome == "Life_satisfaction" ~ "LS",
      outcome == "Negative_experiences" ~ "NE",
      outcome == "Positive_experiences" ~ "PE"
    )
  ) %>% 
  mutate(panel = str_glue("{age} ({sex})")) %>% 
  ggplot(aes(year, val, group = country)) +
  scale_color_brewer(palette = "Dark2") +
  scale_y_continuous(
    "Value", 
    expand = expansion(.01),
    breaks = extended_breaks(),
    labels = ~percent(.)
  ) +
  scale_x_continuous(
    "Year", 
    breaks = seq(2008, 2018, by = 2),
    labels = ~str_replace(., "20", "'")
  ) +
  geom_line(size = .1) +
  facet_grid(
    outcome~panel, 
    scales = "free_y",
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  theme(
    strip.text.x = element_text(margin = margin(2,0,2,0, "pt")),
    strip.text.y = element_text(margin = margin(0,2,0,2, "pt")),
    legend.position = "none",
    text = element_text(size = 5)
  )

@@

Variations in such confounding factors are likely to be significantly smaller within countries over time. We therefore next focused on the model's within-country associations, which measure the extent to which Facebook adoption in a given country predicted well-being in that country adjusting for general linear temporal trends in the well-being outcome. On average across age, and sex, we found that years with greater than the country's average DAU were associated with greater levels of life satisfaction and positive experiences, and lower levels of negative experiences (Table 1), although the first association's 95%CI contained zero. For the average country in our sample, Facebook adoption positively predicted well-being. The magnitudes of these associations were small; a one-percent increase in DAU predicted a `r tab_world[3,4]` increase in positive experiences for the average country. While these associations indicate relations within countries and adjust for confounders that vary linearly with time by including time as a predictor, they are still susceptible to confounders and do not indicate causal relations. Rather, this association describes that, all else being equal, years with greater Facebook adoption tended to be those with greater levels of well-being.

We then turned to our second question and assessed whether within-country associations linking Facebook adoption to well-being differed between age and sex. Differences in the associations were greater between the younger and older demographics than they were between females and males (Figure 2, Average). DAU predicted life satisfaction more positively in the younger age bracket, but these estimates' CIs did not exclude zero. DAU predicted negative experiences negatively only in the younger age group. These differences were less pronounced for positive experiences. These results show that Facebook adoption was, for the average nation in our sample, more positively predictive of well-being for individuals below 35, and that differences between males and females were generally smaller than those between age groups. Importantly, they did not indicate that the well-being of younger individuals is particularly linked to social media adoption.

However, while informative aggregates, these results do not describe associations between Facebook adoption and well-being for any individual country, but rather their averages. To answer our third question, we computed country-specific estimates for each sex and age group (Figure 2). For life satisfaction, `r tmp[1, 5]` countries had credibly positive average associations. For negative experiences, `r tmp[2, 6]` countries had credibly negative average association, and `r tmp[3, 5]` countries had a credibly positive average association between DAU and positive experiences. There were no credible (at the 95% level) country-level negative associations between DAU and life-satisfaction. Although many country-level estimates were uncertain, these results aligned with the general pattern of positive associations between Facebook adoption and country-level well-being.

We found that, overall, Facebook adoption positively predicted well-being. This association held when comparing countries to other countries and to themselves over time. While these descriptive results do not speak to causal effects, they align with other recent findings suggesting that technology has not become increasingly harmful over time [@vuorreThereNoEvidence2021], and that the increased adoption of Internet technologies is not, overall, associated with widespread psychological harms [@vuorreGlobalWellbeingMental2022]. That said, these conclusions are qualified because our focus on daily active users could miss those who use Facebook less regularly. To test this possibility, we also conducted our analyses using monthly active Facebook users (MAU) as the predictor. Those results were substantively the same (see online supplement).

@@

tmp %>%
  ggplot() +
  aes(
    x = Estimate, 
    y = fct_rev(country), 
    col = Age, 
    fill = interaction(Age, Star, sep = " "), 
    shape = Sex
  ) +
  scale_y_discrete(
    "Country",
    expand = expansion(.015),
    labels = ~str_replace(., "World", "Average")
  ) +
  scale_x_continuous(
    "Association with daily active Facebook users (95%CI)",
    breaks = extended_breaks(7),
    expand = expansion(.1)
  ) +
  scale_shape_manual(
    values = c(21, 22)
  ) +
  scale_color_brewer(
    "Age", palette = "Set1", aesthetics = c("color")
  ) +
  scale_fill_manual(values = c("white", "white", "#E41A1C", "#377EB8")) +
  geom_vline(xintercept = 0, lty = 2, size = .25, col = "grey50") +
  geom_linerangeh(
    aes(xmin = CI.Lower, xmax = CI.Upper),
    size = .15, position = position_dodge2v(.7), 
    show.legend = FALSE
  ) +
  geom_point(
    size = .65, stroke = .45,
    position = position_dodge2v(.7)
  ) +
  facet_wrap(
    "outcome", scales = "free_x", nrow = 1,
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(size = 1.5)),
    shape = guide_legend(override.aes = list(size = 1.5))
  ) +
  theme(
    text = element_text(size = 5),
    strip.text.x = element_text(margin = margin(2,0,2,0, "pt")),
    axis.text.y = element_text(size = rel(0.9)),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(
      size = .1, linetype = 2, color = "gray60"
    ),
    legend.position = "bottom", 
    legend.box.margin = margin(-15, 0, -10, 0, "pt"),
    legend.background = element_rect(fill = NA),
    legend.key = element_rect(fill = NA),
    legend.text = element_text(size = rel(0.9))
  )
```

@@

Both datasets (Facebook and Gallup) are proprietary and we therefore could not share them with this manuscript. Our analytic code, along with synthetic datasets, is available at (LINK). Researchers can contact Facebook (ccobb@fb.com) to reproduce our analyses with the actual Facebook adoption dataset. The Gallup data are available online to subscribing institutions.

@@

We studied two metrics of Facebook adoption, daily (DAU) and monthly active users (MAU), from 2008 to 2019 for 72 countries (see Figure 2). FB estimates yearly DAU & MAU with their corresponding means from June 1 to August 31, and rounds values greater than 10,000,000 to three significant digits, and values less than 10,000,000 to the nearest 10,000. We report analyses based on DAU in the main manuscript, and MAU in the online supplement. Facebook calculates DAU and MAU estimates separately for individuals aged 13-34 and 35+. User age is determined based on Facebook profile information, which can be inaccurate (e.g. young users reporting an older age.) Accordingly, Facebook has trimmed 0.008% of total MAU to exclude accounts with unrealistic or non-reported ages.

@@

Facebook defines users as individuals who visited Facebook or used Messenger during the appropriate period. Although accurately captured, these numbers are not perfect indicators of actual user numbers because of possible duplicate and false accounts. Facebook estimates those accounts to account for 11% and 5% of global MAUs, respectively, and that the former may be more likely in developing regions. Because internal criteria and methodology for determining duplicate and false accounts can change over time, estimates of MAU and DAU can also change.

@@

$$
\begin{aligned} 
y_i &\sim \text{Normal}(\mu_i, \sigma^{y2}v_i^2), \\
\mu_i &= \alpha_{0} + \beta_{0\text{country}[i]} + \\
&\quad\ (\alpha_{1} + \beta_{1\text{country}[i]})\text{Sex}_i + \\
&\quad\ (\alpha_{2} + \beta_{2\text{country}[i]})\text{Age}_i + \\
&\quad\ (\alpha_{3} + \beta_{3\text{country}[i]})\text{Time}_i + \\
&\quad\ (\alpha_{4} + \beta_{4\text{country}[i]})\text{DAU}^{\text{CW}}_i + \\
&\quad\ (\alpha_{5} + \beta_{5\text{country}[i]})\text{Sex}_i \times \text{Time}_i + \\
&\quad\ (\alpha_{6} + \beta_{6\text{country}[i]})\text{Age}_i \times \text{Time}_i + \\
&\quad\ (\alpha_{7} + \beta_{7\text{country}[i]})\text{Sex}_i \times \text{Age}_i + \\
&\quad\ (\alpha_{8} + \beta_{8\text{country}[i]})\text{Sex}_i \times \text{Age}_i \times \text{Time}_i + \\
&\quad\ (\alpha_{9} + \beta_{9\text{country}[i]})\text{Sex}_i \times \text{DAU}^{\text{CW}}_i + \\
&\quad\ (\alpha_{10} + \beta_{10\text{country}[i]})\text{Age}_i \times \text{DAU}^{\text{CW}}_i + \\
&\quad\ (\alpha_{11} + \beta_{11\text{country}[i]})\text{Sex}_i \times \text{Age}_i \times \text{DAU}^{\text{CW}}_i + \\
&\quad\ \alpha_{12} \text{DAU}^{\text{CB}}_i,
\\
\pmb{\beta} &\sim \text{MVN}(\pmb{0}, \Sigma).
\end{aligned}
$$

Where $y_i$ is the outcome (e.g. life satisfaction) on row $i$, $v_i$ its standard error, $\pmb{\alpha}$ are the population-level coefficients, $\beta_{0, \dots, 11\text{country}[i]}$ are the country-specific coefficients for the country indicated on row $i$, $\text{DAU}^{\text{CW}}_i$ the within-country centered year-aggregated daily (or monthly) active Facebook users, and $\text{DAU}^{\text{CB}}_i$ the between-country centered DAU (or MAU).

@@

tmp %>% 
  ggplot() +
  aes(
    x = Estimate, 
    y = fct_rev(country), 
    col = Age, 
    fill = interaction(Age, Star, sep = " "), 
    shape = Sex
  ) +
  scale_y_discrete(
    "Country",
    expand = expansion(.015),
    labels = ~str_replace(., "World", "Average")
  ) +
  scale_x_continuous(
    "Association with monthly active Facebook users (95%CI)",
    breaks = extended_breaks(7),
    expand = expansion(.1)
  ) +
  scale_shape_manual(
    values = c(21, 22)
  ) +
  scale_color_brewer(
    "Age", palette = "Set1", aesthetics = c("color")
  ) +
  scale_fill_manual(values = c("white", "white", "#E41A1C", "#377EB8")) +
  geom_vline(xintercept = 0, lty = 2, size = .25, col = "grey50") +
  geom_linerangeh(
    aes(xmin = CI.Lower, xmax = CI.Upper),
    size = .15, position = position_dodge2v(.7), 
    show.legend = FALSE
  ) +
  geom_point(
    size = .65, stroke = .45,
    position = position_dodge2v(.7)
  ) +
  facet_wrap(
    "outcome", scales = "free_x", nrow = 1,
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(size = 1.5)),
    shape = guide_legend(override.aes = list(size = 1.5))
  ) +
  theme(
    text = element_text(size = 5),
    strip.text.x = element_text(margin = margin(2,0,2,0, "pt")),
    axis.text.y = element_text(size = rel(0.9)),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(
      size = .1, linetype = 2, color = "gray60"
    ),
    legend.position = "bottom", 
    legend.box.margin = margin(-15, 0, -10, 0, "pt"),
    legend.background = element_rect(fill = NA),
    legend.key = element_rect(fill = NA),
    legend.text = element_text(size = rel(0.9))
  )
```
